Hydraulically-vulnerable trees are buffered by deep-water access in a tropical forest community

### 2020-10-09 manuscript.rmd
indexed an equation
stream discharge details

\
We calibrated ELM-FATES over 2012-2018 against three key fluxes and states in Eq. \@ref(eq:water-balance), namely, (i) evapotranspiration ($ET \simeq E_{v} + E_{g}$) from the eddy flux tower (2012-2017; see [SI](#SIData.matteo)), (ii) stream discharge for the Conrad Catchment ($q_{run} \simeq q_{over} + q_{drai}$; 2012-2018) and (iii) soil volumetric water content (VWC, $\theta_{z} \simeq W_{liq, z}$) from two sources: a long-term (2012-2018) record of VWC from a point location near the eddy flux tower we built with a vertical TDR probe over 0-15 cm depth ($\theta_{0-0.15 m}$) and existing 50 ha plot-wide snap-shot measurements of VWC during the dry season of 2015 and 2016 for the depths of 0.15, 0.4 and 1 m (1299 samples in total covering all soil types and habitats; see [SI](#SIData.VWC) and @Kupers:2019 for details; hereafter, $\theta_{0.15 m}$, $\theta_{0.4m}$ and $\theta_{1 m}$ refer to plot-wide averages).

@@

\begin{equation}
K_{leaf,~s} = A_{s}e^{-B_{s}\Psi_{leaf}} + error
(\#eq:vc-exp)
\end{equation}
where $A$ and $B$ are fitted species-specific parameters. From these fits we estimated species $K_{max,~leaf}$, $\Psi_{50,~leaf}$ and  $\Psi_{80,~leaf}$ (see Table 1). We identified scaling relationships between fitted parameters $A$ and $B$ in Eq. \@ref(eq:vc-exp) and two traits that were available at the community-scale, namely,  WSG and LMA (@Wright:2010; see SI for description of how data gaps were filled. We fitted polynomial equations described as,

@@

@@

```{r gpp-vpd, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', dpi = 96, out.width = "35%", fig.pos = 'H', results = "asis", fig.cap = "\\label{lab-gpp-vpd}Scatterplot of mean daily gross primary productivity (GPP) and mean daily vapor pressure deficit (VPD) by month measured at the eddy flux tower in Barro Colorado Island over 2012-2017. Red line shows a fit from a polynomial model described in inset, and used to standardize VPD employed in equations \\@ref(eq:M2) and \\@ref(eq:M4)."}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/gpp_by_env_variables/monthly_gpp.vpd.jpeg"))
```

>>>>

\
We calibrated ELM-FATES over 2012-2018 against three key fluxes and states in Eq. \@ref(eq:water-balance), namely, (i) evapotranspiration ($ET \simeq E_{v} + E_{g}$) from the eddy flux tower (2012-2017; see [SI](#SIData.matteo)), (ii) stream discharge from the Conrad catchment ($q_{run} \simeq q_{over} + q_{drai}$; 2012-2018; see [SI](#SIData.steveP)) and (iii) soil volumetric water content (VWC, $\theta_{z} \simeq W_{liq, z}$) from two sources: a long-term (2012-2018) record of VWC from a point location near the eddy flux tower we built with a vertical TDR probe over 0-15 cm depth ($\theta_{0-0.15 m}$) and existing 50 ha plot-wide snap-shot measurements of VWC during the dry season of 2015 and 2016 for the depths of 0.15, 0.4 and 1 m (1299 samples in total covering all soil types and habitats; see [SI](#SIData.VWC) and @Kupers:2019 for details; hereafter, $\theta_{0.15 m}$, $\theta_{0.4m}$ and $\theta_{1 m}$ refer to plot-wide averages).

@@

\begin{equation}
K_{leaf,~s} = A_{s}e^{-B_{s}\Psi_{leaf_{s}}} + error
(\#eq:vc-exp)
\end{equation}
where $A$ and $B$ are fitted species-specific parameters. From these fits we estimated species $K_{max,~leaf}$, $\Psi_{50,~leaf}$ and  $\Psi_{80,~leaf}$ (see Table 1). We identified scaling relationships between fitted parameters $A$ and $B$ in Eq. \@ref(eq:vc-exp) and two traits that were available at the community-scale, namely,  WSG and LMA (@Wright:2010; see SI for description of how data gaps were filled. We fitted polynomial equations described as,

@@

### Stream discharge from the Conrad catchment (#SIData.steveP)

The Conrad weir is located on the gently sloping western side of Barro Colorado Island. The weir consists of a 90 degree ‘V’ notch with a rectangular upper section. The weir was constructed in several stages between 1993 and 1996. Stream stage is recorded at 5-minute intervals.

The weir drains a catchment of approximately 40 ha of which at least 90% consists of the central plateau area of the island. Peak flow usually follows peak rain by approximately 2-3 hours. Large storms are sometimes preceded by small, initial peaks generated by the rain falling on the steeper slopes close to the weir.

Streamflow data are analyzed by software custom-built by author SP. The program is used to correct for blockages of the weir ‘V’ by fallen vegetation (a common occurrence), sensor drift, small gaps and other problems. Discharge is then calculated using an empirically derived rating curve calculated for each 1mm stream stage.

@@

```{r gpp-vpd, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', dpi = 96, out.width = "35%", fig.pos = 'H', results = "asis", fig.cap = "\\label{lab-gpp-vpd}Scatterplot of mean daily gross primary productivity (GPP) and mean daily vapor pressure deficit (VPD) by month measured at the eddy flux tower in Barro Colorado Island over 2012-2017. Red line shows a fit from a polynomial model described in inset, and used to standardize VPD employed in equations \\@ref(eq:M2) and \\@ref(eq:M4)."}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/gpp_by_env_variables/daily_gpp.vpd.jpeg"))
```
